---
title: "Proyect 1"
author: "Brayan Ortiz"
date: "5/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, let's load the data:

```{r}
if (!file.exists("activity.csv")) unzip("activity.zip")
act <- read.csv("activity.csv")
```

1. What is mean total number of steps taken per day?

```{r}
totalday <- tapply(act$steps, act$date, sum, na.rm=TRUE)
hist(totalday)
meanday <- tapply(act$steps, act$date, mean, na.rm=TRUE)
medianday <- tapply(act$steps, act$date, median, na.rm=TRUE)
meanday
medianday
```

2. What is the average daily activity pattern?

```{r}
y <- tapply(act$steps, act$interval, mean, na.rm=TRUE)
plot(unique(act$interval), y, type="l", xlab="5-minute interval", ylab="Average number of steps")
a <- names(y[which.max(y)])
```

The 5-minute interval with the maximun numbers of steps is the interval `r a`.

3. Imputing missing values


```{r}
act2 <- act
datamissing <- act2[is.na(act$steps),]
nmissing <- nrow(datamissing)
for (i in 1:nmissing) {
    datamissing[i,"steps"] <- y[as.character(datamissing[i,"interval"])]
}
act2[is.na(act2$steps),"steps"] <- datamissing$steps
```

The total number of missing values is `r nmissing`.  
For filling in all of the missing values it was used the mean for that 5-minute interval.

```{r}
totalday2 <- tapply(act2$steps, act2$date, sum)
hist(totalday2)
meanday2 <- tapply(act2$steps, act2$date, mean)
medianday2 <- tapply(act2$steps, act2$date, median)
meanday2
medianday2
```

These values differ from the estimates from the first part of the assignment, but the difference is small. Filling in the missing values results in greater values for the total, mean, and median steps in some days. Specially for the days with a greater proportion of missing values.

4. Are there differences in activity patterns between weekdays and weekends?

```{r}
act2$day <- weekdays(as.Date(act2$date))
act2$week <- as.factor(ifelse(act2$day=="sábado" | act2$day=="domingo", "weekend", "weekday"))
two <- split(act2, act2$week)
daymean <- tapply(two$weekday$steps, two$weekday$interval, mean)
endmean <- tapply(two$weekend$steps, two$weekend$interval, mean)
par(mfrow=c(1,2))
plot(unique(act2$interval), daymean, type="l", main="Weekdays", xlab="5-minute interval", ylab="Average number of steps", ylim=c(0,230))
plot(unique(act2$interval), endmean, type="l", main="Weekends", xlab="5-minute interval", ylab="Average number of steps", ylim=c(0,230))
```

